Robust Estimation, Prediction and Control with \\ Linear Dynamics and Generic Costs

### 2020-02-03 main.tex
Update

Despite the dazzling successes of Reinforcement learning in cracking many problems considered unsolvable in the past decade [ref needed], this tool has hardly been applied in real industrial issues. This could be attributed to two undesirable properties which limit its practical applications: the dependence on a tremendous amount of interaction data that cannot always be simulated, and the reliance on trial-and-error and random exploration. In this paper, we consider the problem of controlling an unknown linear system $x(t)$ so as to maximise an \emph{arbitrary} bounded reward function $R$, in a critical setting where mistakes are costly and must be avoided at all time. 
This choice of the rich reward space is crucial to have sufficient flexibility to model non-convex and non-smooth functions that naturally arise in many practical problems such as combinatorial optimisation, obstacle avoidance, etc.

@@

To address this issue, we turn to the framework of \emph{robust} decision-making: instead of merely considering a point estimate of the dynamics, we build an entire \emph{confidence region} that contains the true dynamics with high probability, and optimise the decisions so as to maximise the worst-case outcome with respect to this uncertainty. Minimax problem are hard to solve, especially since the reward function $R$ can be arbitrary. In particular, this keeps us from employing traditional tools from the robust synthesis literature, such as $\cH_\infty$ \todo{(ref needed)} control or system-level synthesis \todo{(ref needed)}. Since we are optimising a generic reward function, we turn to more general sequential optimisation algorithms. The continuous state space prevents us from using dynamic programming techniques, so we consider tree-based planning instead. Though some work address continuous actions \todo{ref needed}, we resort to a first approximation and discretise the decision space by adopting a hierarchical control architecture. We  formulate the problem high-level decision of selecting a low-level controller among a finite basis \todo{(ref from temporal abstraction)}. For instance, this could amount to choosing a short-term goal to follow, or a skill to execute such as a driving maneuver. The passage from continuous to discrete optimization induces a suboptimality, but it can be mitigated by diversifying the controller basis. These low-level controllers are not learnt, which also constitutes a limitation \todo{(argue it is realistic in industrial setting?)}. However, tree-based planning algorithms are only designed for a single known generative model instead of an entire confidence region for the system dynamics. Hence, we adapt them to the robust framework by following a two-steps procedure: we first predict a set of states that can be reached under a planned control sequence with respect to all possible dynamics. This reachability set is used to provide a conservative pessimistic reward to the planning algorithm.

@@

The dynamics of the real system are described in continuous time. However, all the sensing and control performed on-board happen in discrete time, with time-step $\dd t$. For any variable $z$, we use the subscript notation to refer to these discrete times: $z_n = z(t_n)$ with $t_n = n\dd t$.

@@

\begin{algorithm}[tb]
	\caption{Robust Estimation, Prediction and Control}
	\label{alg:full}
	\begin{algorithmic}
		\STATE {\bfseries Input:} confidence level $\delta$
		\FOR{$N = 1,2,3,\dots$} 
		\STATE $\cC_{[N],\delta} \gets$\textsc{Model Estimation}$(\cD_{[N]})$. \eqref{eq:polytope}
		\FOR{planning steps $k\in[K]$}
		\FOR{$a\in A$}
		\STATE $[\underline{x}_{n+1}, \overline{x}_{n+1}]\gets$ \textsc{Interval Prediction}(
		\STATE $\qquad\qquad\qquad\qquad\cC_{[N],\delta}, a_k$). \eqref{eq:interval-predictor}
		\ENDFOR
		\STATE $a_{k+1}$ $\gets$\textsc{Pessimistic Planning}$($
		\STATE$\qquad\qquad\underline{R_{n+1}}([\underline{x}_{n+1}, \overline{x}_{n+1}]))$.  \eqref{eq:opd}
		\ENDFOR
		\STATE Execute the first recommended control $u_N$, and add the observed transition $(x_N, y_N)$ to $\cD_{[N+1]}$.
		\ENDFOR
	\end{algorithmic}
\end{algorithm}

@@

\paragraph{LQ systems} There is a rich body of literature dedicated to the study of LQ systems, and we focus on those providing non-asymptotic results when the system parameters $A, B$ are unknown. In their seminal work \citet{abbasi-yadkori11a}, consider the problem of cumulative regret minimisation. They follow the \emph{Optimism in the Face of Uncertainty} paradigm from the multi-armed bandit literature, that consists in selecting an optimistic system within a high-confidence region while enforcing a controllability constraint, and computing the corresponding optimal control in closed-form by solving a Riccati equation. They show that this procedures achieves a $\tilde{\cO}\left(T^{1/2}\right)$ cumulative regret. \citet{abeille18a} provide the same result by applying Thomson Sampling rather of optimism to select a dynamical model, and rejection sampling to enforce the controllability constraint.
While previous methods suffered from limited applicability and computationally intractable subroutines, \citep{Dean2018} provide a polynomial-time algorithm which achieves a suboptimal cumulative regret of $\tilde{\cO}\left(T^{2/3}\right)$.

Offline estimation and control synthesis: \citep{Dean2017}. Summary: Estimation procedure of $N$ episodes of duration $T$ with gaussian excitation, ellipsoid in L2 norm, robust controller with System Level Synthesis, simple regret in $O(\sqrt{\frac{\log(1/\delta)}{N}})$

We wish to act straight away in a reasonable (structure) but cautious (robust) way, and progressively get more and more aggressive as we obtain more data and reduce uncertainty.

Thus, we do not wish to explore actively the more promising regions as in OFU, nor the more likely region as in TS, nor act randomly as in Coarse-ID They inject exploration noise.

We have linear dynamics but our cost is not quadratic. Quadratic cost: stabilize a system around a reference trajectory known a priori. Thus, we can accurately represent planning tasks with discrete/branching decisions involving discontinuities (e.g. collision vs non-collision states).

>>>>

Despite the dazzling successes of Reinforcement learning in cracking many problems considered unsolvable in the past decade [ref needed], this tool has hardly been applied in real industrial issues. This could be attributed to two undesirable properties which limit its practical applications: the dependence on a tremendous amount of interaction data that cannot always be simulated, and the reliance on trial-and-error and random exploration. In this paper, we consider the problem of controlling an unknown linear system $x(t)$ so as to maximise an \emph{arbitrary} bounded reward function $R$, in a critical setting where mistakes are costly and must be avoided at all time. 
This choice of rich reward space is crucial to have sufficient flexibility to model non-convex and non-smooth functions that naturally arise in many practical problems such as combinatorial optimisation, obstacle avoidance, etc., while quadratic costs are only suited for the tracking of a reference trajectory known a priori.

@@

To address this issue, we turn to the framework of \emph{robust} decision-making: instead of merely considering a point estimate of the dynamics, we build an entire \emph{confidence region} that contains the true dynamics with high probability, and optimise the decisions so as to maximise the worst-case outcome with respect to this uncertainty. Minimax problem are hard to solve, and all the more in our setting where the reward function $R$ is generic. In particular, this excludes a number of tools from the robust synthesis literature, such as $\cH_\infty$ \todo{(ref needed)} control or system-level synthesis \todo{(ref needed)}. We turn instead to more general planning algorithms. The continuous state space prevents us from using dynamic programming techniques, so we consider tree-based planning instead. Though some work address continuous actions \todo{ref needed}, we resort to a first approximation and discretise the decision space by adopting a hierarchical control architecture. We  formulate the problem high-level decision of selecting a low-level controller among a finite basis \todo{(ref from temporal abstraction)}. For instance, this could amount to choosing a short-term goal to follow, or a skill to execute such as a driving maneuver. The passage from continuous to discrete optimization induces a suboptimality, but it can be mitigated by diversifying the controller basis. However, tree-based planning algorithms are only designed for a single known generative model instead of an entire confidence region for the system dynamics. Hence, we adapt them to the robust framework by following a two-steps procedure: we first predict a set of states that can be reached under a planned control sequence with respect to all possible dynamics. This reachability set is used to provide a conservative pessimistic reward to the planning algorithm. The next section aims to clarify and formalize these ideas.

@@

The dynamics of the real system are described in continuous time. However, all the sensing and control performed on-board happen in discrete time, with time-step $\dd t$. For any variable $z$, we use the subscript notation to refer to these discrete times: $z_n = z(t_n)$ with $t_n = n\dd t$. We use bold symbols to denote sequences $\bz = (z_1,z_2,\dots)$.

@@

\begin{algorithm}[tb]
	\caption{Robust Estimation, Prediction and Control}
	\label{alg:full}
	\begin{algorithmic}
		\STATE {\bfseries Input:} confidence level $\delta$
		\FOR{$N = 1,2,3,\dots$} 
		\STATE $\cC_{[N],\delta} \gets$\textsc{Model Estimation}$(\cD_{[N]})$. \eqref{eq:polytope}
		\FOR{planning steps $k\in[K]$}
		\FOR{$b\in A$}
		\STATE $[\underline{x}_{n+1}, \overline{x}_{n+1}]\gets$ \textsc{Interval Prediction}(
		\STATE $\qquad\qquad\qquad\qquad\cC_{[N],\delta}, a_kb$). \eqref{eq:interval-predictor}
		\ENDFOR
		\STATE $a_{k+1}$ $\gets$\textsc{Pessimistic Planning}$($
		\STATE$\qquad\qquad\underline{R_{n+1}}([\underline{x}_{n+1}, \overline{x}_{n+1}]))$.  \eqref{eq:opd}
		\ENDFOR
		\STATE Execute the first recommended control $u_N$, and add the observed transition $(x_N, y_N)$ to $\cD_{[N+1]}$.
		\ENDFOR
	\end{algorithmic}
\end{algorithm}

@@

\paragraph{Linear Quadratic systems} There is a rich body of literature dedicated to the study of LQ systems, and we focus on those providing non-asymptotic results when the system parameters $A, B$ are unknown. In their seminal work \citet{abbasi-yadkori11a}, consider the problem of cumulative regret minimisation. They follow the \emph{Optimism in the Face of Uncertainty} paradigm from the multi-armed bandit literature, that consists in selecting an optimistic system within a high-confidence region while enforcing a controllability constraint, and computing the corresponding optimal control in closed-form by solving a Riccati equation. They show that this procedures achieves a $\tilde{\cO}\left(T^{1/2}\right)$ cumulative regret. \citet{abeille18a} provide the same result by applying Thomson Sampling rather of optimism to select a dynamical model, and rejection sampling to enforce the controllability constraint. These methods which aim at regret minimization are actively exploring to quickly reduce uncertainty, through information-seeking behaviours. Other work use noise injection for exploration such as \citep{Dean2017} or \citep{Dean2018}, who provide a polynomial-time algorithm with a cumulative regret of $\tilde{\cO}\left(T^{2/3}\right)$. In contrast, our main objective is to prevent failures, which forbids to explore uncertain regions.
% They also suffer from a limited applicability and computationally intractable subroutines.
%$
The work of \citet{Dean2017} is closest to our setting: after an offline estimation phase under random exploration, they estimate a simple regret between a robust controller and the oracle optimal performance.
